---
title: "Practical_maximum_likelihood_estimation_ Oscar_Contreras_Rafael_Castilla"
author: "F.R.Castilla, O.Contreras"
date: "2022-10-03"
output: 
  pdf_document:
      latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## R Markdown

Resolve the following exercise in groups of two students. Write your solution in a Word, Latex or Markdown document and generate a pdf file with your solution. 
Upload the pdf file with your solution to the corresponding task at the Moodle environment of the course, no later than the hand-in date.

1.(16p) ML estimation of a one-parameter distribution. 
  Let X be a random variable with probability density
$f(x|\beta)=\beta x^{b-1}$ with $\geq x\geq1$, $\beta >0$
 we consider a random sample of n observation of this distribution.
 
 a) (2p) Write down the likelihood function for a sample of n observations of this distribution.
 <div style="background-color:#F0F0F0">
&emsp;<i class="fas fa-comment-dots"></i>
Answer: $L(f(x|\beta))=\prod_{i=1}^{n} f(x|\beta) =\prod_{i=1}^{n} \beta x^{b-1}_{i}=\beta^{n}\prod_{i=1}^{n}x_i^{b-1}$ 
&emsp;
</div>
```{r}
like<-function(x,b){
  prod<-1
  for (i in 1:len(x)){
   prod<-prod*(x[i]^{b-1})
  }
  l<-b^len(x)*prod
  return(l)
}
```
b) (1p) Obtain the log-likelihood function
$l(L(f(x|\beta)) = nlog(\beta)+\sum^n_{i=1}(\beta-1)log(x_i)=nlog(\beta)+n(\beta-1)\sum^{n}_{i-1}log(x_i)$
```{r}
loglike<-function(x,b){
  sum<-0
  for(i in 1:len(x)){
    sum<-sum+log(x[i])
  }
  l<-n*log(b)+n(b-1)*sum
  return(l)
}
```

c) (2p) Find the stationary point(s) of the log-likelihood function analytically.
$\frac{dl(x|\beta)}{d\beta}=\frac{n}{\beta}$
for find the stationary poitt the derivate qual to 0
$0=\frac{n}{\beta}$


d)(1p) Determine whether the stationary point(s) are maxima or minima.

for know if the stationary point is maxima or minima the need the 2 derivate
$\frac{d^2l}{d^2\beta}=-\frac{n}{x^2}$
and if the result is positive is the minimum and if is negative is a maximum

```{r}
x<-read.table('~/Statistic/sample.dat')
```
            
            