---
title: "Practical_maximum_likelihood_estimation_ Oscar_Contreras_Rafael_Castilla"
author: "F.R.Castilla, O.Contreras"
date: "2022-10-03"
output: 
  pdf_document:
      latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## R Markdown

Resolve the following exercise in groups of two students. Write your solution in a Word, Latex or Markdown document and generate a pdf file with your solution. 
Upload the pdf file with your solution to the corresponding task at the Moodle environment of the course, no later than the hand-in date.

1.(16p) ML estimation of a one-parameter distribution. 
  Let X be a random variable with probability density
$f(x|\beta)=\beta x^{b-1}$ with $\geq x\geq1$, $\beta >0$
 we consider a random sample of n observation of this distribution.
 
 a) (2p) Write down the likelihood function for a sample of n observations of this distribution.
 <div style="background-color:#F0F0F0">
&emsp;<i class="fas fa-comment-dots"></i>
Answer: $L(f(x|\beta))=\prod_{i=1}^{n} f(x|\beta) =\prod_{i=1}^{n} \beta x^{b-1}=\beta^{n}x^{nb-n}$ 
&emsp;
</div>
```{r}
likelihood<-function(x,b,n){
  l=b^n*x^(n*b-n)
  return(l)
}
```
b) (1p) Obtain the log-likelihood function
