---
title: "Likelihood ratio tests"
author: "F.R.Castilla"
date: "2022-10-17"
output: 
  pdf_document:
      latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Resolve the following exercises in groups of two students. Write your solution in a Word, Latex or Markdown document and generate a pdf file with your solution. Upload the pdf file with your solution to the corresponding task at the Moodle environment of the course, no later than the hand-in date.


### Many well-known standard statistical tests are actually LRT tests. We do some exercises with data sets where we apply these LRT tests.

1. (10p) Likelihood 
ratio test for Hardy-Weinberg equilibrium. In a genetic association study, the genotypes of a single nucleotide polymorphism have been determined for a sample of individuals. The genotype data file snp.dat contains the genotyping results.

(a) (1p) Load the data in the R environment, and make a table of the different genotypes. Report the table. What is the sample size of the study?
```{r}
genotype<-read.table("snp.dat")
table(genotype)
n=nrow(genotype)
print(paste0("the sample size of this study is: ",n))
```


(b) (1p) How many alleles does this SNP have? How many geno-
types could it theoretically have? Estimate all relative genotype
frequencies by maximum likelihood (ML). Report the values of
the ML estimators.

have 2 genotype and 3 fenotype

```{r}
#the frequence
x<-c(AT=40,TT=100)
p_AT=x[1]/n
p_AT
p_TT=x[2]/n
p_TT

```


(c) (2p) Count the number of alleles of each type in the sample.
Estimate the relative allele frequencies by ML. Report the values
of the ML estimators.
```{r}
A=(0+40)/(140*2)
T=(40+100*2)/(140*2)
A
T


```

(d) (1p) Which allele is the minor (least common) allele?

the allele minor is A


(e) (1p) Do a likelihood ratio test (LRT) for Hardy-Weinberg equilibrium using the HWLratio function of the R-package HardyWein-
berg. Report the likelihood ratio statistic and the p-value.
```{r}
library(HardyWeinberg)
x <- c(AA=0,AT=40,TT=100)
HWLratio(x,verbose=TRUE)
``` 
(f) (1p) State your conclusion of the LRT.
p-value is less that 0.05 and i reject the neutral hypothesis and the data not follow the hardy-winberg equilibrium

(h) (1p) Calculate the p-value ”by hand” using the value observed for the LR statistic and its distribution. Show your computations.
Do you obtain the same result as the HWLratio function?
$x^2=n(\frac{4n_{AA} . n_{TT}-n_{AT}^2}{(2n_{AA}+n_{AT})(2n_{TT}+n_{AT})})²$
```{r}
n
x <- c(AA=0,AT=40,TT=100)
l1<-(4*x[1]*x[3]-x[2]^2)
l1
l2<-(2*x[1]+x[2])*(2*x[3]+x[2])
l2

X2<-n*(l1/l2)^2
X2
pchisq(X2, 1, lower.tail = TRUE)
```

